---
title: "Multispecies Loops"
author: "Kelsey King"
date: '2022-05-27'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(ggpubr)
```

# Data

This data comes from GBIF.org (27 May 2022) GBIF Occurrence Download
<https://doi.org/10.15468/dl.vf6fgm>. We collected all lupines
(*Lupinus* spp.) in the United States, with coordinate uncertainty less
than \~5000 meters. If you utilize this data elsewhere, please use the
above citation, see [GBIF citation
guidelines](https://www.gbif.org/citation-guidelines).

Now let's get started. You'll first want to import the data. I cleaned
the data as well to give a smaller set to work with, the original data
is included in the zip folder, but the file we use below is the cleaned
data, just what we want to use in the tutorial.

Data was filtered to records identified to species or finer, states with
more than 80 records, and records that have county information. Then we
removed records with missing date information.

```{r}
data=read.csv("tutorial_data.csv")

#forming a , which combo name
#could be used for various sorting
#I use this as a unique for the type of sort I want
#could be used to make it only one loop instead of nested/
data$sppState=as.character(paste(data$specificEpithet,
                          data$stateProvince))

#for my figures I only want to display 
#one point for each date in each county,
#for each species
#so I can use this code.
data$day.loc=as.character(paste(data$county, 
    data$specificEpithet,
    data$year,data$startDayOfYear))

#removing duplicates
data1=distinct(data, day.loc, .keep_all = TRUE)

#sorting to more than 50 records for the species-state combo name
summary.dat=data1 %>% group_by(sppState) %>%
  summarize(sample.size=n())
keep.dat=summary.dat[summary.dat$sample.size>50,]

#finally altering the data
data2=data1[which(data1$sppState %in% keep.dat$sppState),]

```

# Loop

```{r}
species=unique(data2$specificEpithet)
#loop for each species
for(each.spp in 1:length(species)){
  
  #set the current species
  cur.spp=species[each.spp]
  
  #sort to the current species data
  spp.dat=data2[data2$specificEpithet==cur.spp,]
  
  #make a list of sites for that species
  spp.states=unique(spp.dat$stateProvince)
  
  
  #loop for each site
  for(each.loc in 1:length(spp.states)){
    
    #specifying the current site
    cur.state=spp.states[each.loc]
    
    #soting the data to the current site
    cur.dat=spp.dat[spp.dat$stateProvince==cur.state,]
    
    ######################################################
    #here are examples of things you can do inside this loop
    #use this code to guide you on making your own loop
    
        #plotting the day of year for sightings
    ggplot(data=cur.dat,aes(x=year,y=startDayOfYear, 
        color = county))+
      
    #this line of code is what lets add in the 
    #name of the species and the site to the figure
      geom_jitter()+labs(title=paste("Lupinus",cur.spp,"in", 
          cur.state))+xlab("Year")+ylab("Day of Year")+
      
      #pick a theme you like
      #I want to hide the legend
      theme_pubr()+
      theme(legend.position = "none")
      
      #save all the files automatically, with labelling
    ggsave(paste0("./Figures/Lupinus_",cur.spp,"_", 
      cur.state,".jpg"), dpi=300,height=4,width=4)
    }
}

```
